A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing

### 2012-09-08 bidir_pathtrace_skeleton.tex
Quelques corrections sur l'article, vire quelques parties inutiles.

In the 3d cubical complex framework, objects are no more made of voxel, but of cubes, squares, lines and vertices $\mathbb{R}$. 
Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$. Any subset $f$ of $\Z^n$ such that $f$ is the cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$. A $0$-face is called a \myem{vertex}, a $1$-face is an \myem{edge}, a $2$-face is a \myem{square}, and a $3$-face is a \myem{cube}.

@@

Informally, in 3d, a set of face is a cubical complex if, for each cube of the complex, the six squares (sides) of the cube also belong to the complex, for each square of the complex, the four edges (sides) of the square also belong to the complex, and for each edge of the complex, the two vertices (extremities) of the edge also belong to the complex.

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. We denote by $\princ{X}$ the set composed of all facets of $X$ . A complex $X$ is \myem{pure} if all its facets have the same dimension. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex. The notions of purity and dimensions can be trivially extended to sets of faces.

In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$. In $\allfaces{3}$, a complex is thin if it contains no cube.

@@

Traditionally, a binary image is defined as a finite subset of $\Zset^n$ (with $n=2$ or $n=3$). Given $S \subseteq \Zset^n$, the object voxels are the elements of $S$. This kind of image is the most common one in the field of image processing so, in order to work in cubical complex framework, we need to find a way to transpose a binary image to cubical complex framework.

Informally, to do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$ (to a pixel we associate a square, to a voxel we associate a cube). More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$. Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$ (see Fig.~\ref{fig::voxel_to_complex}).

In the following, most of the objects we consider were indeed binary images which were then transposed into cubical complex framework: this is why most of the two-dimensional complexes we show are pure 2-complexes, and most of the three-dimensional complexes we show are pure 3-complexes.

@@

In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$ ($g$ is included only in $f$). It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.

@@

We insist here on the fact that, if $Y$ is a collapse of a complex $X$, then $Y$ is a complex.

Let us now introduce some elements that will serve later for proving the thinness of our skeletons.
Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
(f_0,...,f_\ell)$ of faces of $\allfaces{n}$ such that for all $i \in \dinterval{0}{\ell}$, either $i$ is even and $f_i$ is an $n$-face, or $i$ is odd and $f_i$ is an $(n-1)$-face with $\contstrict{f_i}=\{f_{i-1}, f_{i+1}\}$ (such path always exists).

@@

As in the DT framework, it is sometimes necessary to perform collapse in the cubical complex framework while preserving some faces safe from deletion: these faces are the so-called inhibitor set. When using an inhibitor set during collapse, the guarantee of having a thin result does no more hold. As illustrated on figure ffjkz, the inhibitor set can be thin and a result of the constrained thinning can still not be thin.
However, under some conditions, it is possible to use an inhibitor set while collapsing, and still have the guarantee of getting a thin result. In order to prove this, let us state the following:

\begin{lemma}
\label{lem::remove_small_pairs}
Let $\complex{n}{X}$, let $(f_1,g_1)$ be a free pair of $X$, and $(f_2,g_2)$ be a free pair of $(X \backslash \{f_1,g_1\})$. If $\dim(f_2) > \dim(f_1)$, then $(f_2,g_2)$ is free for $X$. 
\end{lemma}

\begin{proof} If $(f_2,g_2)$ is a free pair of $(X \backslash \{f_1,g_1\})$, then $g_2$ is included in only one face of $(X \backslash \{f_1,g_1\})$, which is $f_2$. As $(\dim(g_2) = \dim(f_2) -1)$ and $(\dim(f_1) \leq (\dim(f_2)-1))$, then $g_2 \nsubseteq f_1$ and $g_2 \nsubseteq g_1$. Therefore, $g_2$ is included in only one face of $X$, which is $f_2$: therefore, $(f_2,g_2)$ is free for $X$.$\finproof$
\end{proof}

This lemma implies that, when one has a sequence of removal of free pairs of faces from a complex, one can only keep the free pairs of highest dimension and still have a sequence of removal of free pairs from the complex.

The following proposition explains which conditions on the inhibitor set (denoted $W$) need to be matched in order to guarantee the thinness of the result of a thinning constrained by $W$.

\begin{proposition}
\label{prop::thin_with_inhibit}
Let $X$ be an $n$-complex with $n>0$, let $S$ be a collapse of $X$ such that $\dim(S) \leq (n-1)$ and let $\subcomplex{W}{S}$. Let $Y$ be a collapse of $X$ such that $W \subseteq Y$ and such that there are no free pairs in $Y$ included in $Y \backslash W$. Then, the dimension of $Y$ is inferior or equal to $(n-1)$.
\end{proposition}

\begin{proof}
In the following, we show that if there exists an $n$-face in $Y$, then it belongs to a free pair for $Y$ that is included in $Y \backslash W$, a contradiction with the hypothesis of the proposition.

Let $C = ((a_1, b_1),...,(a_k, b_k))$ be a sequence of removal of free faces which allows to obtain $S$ from $X$: for all $i \in [1;k], (a_i, b_i)$ is free for $X \backslash \{a_1, b_1, ..., a_{i-1}, b_{i-1}\}$ and $S = X \backslash \{a_1, b_1, ..., a_k, b_k\}$. 
Let $C'$ be the sequence $C$ restrained only to free pairs containing an $n$-face: $C' = ((f_1, g_1),...,(f_h, g_h))$. A consequence of lemma \ref{lem::remove_small_pairs} is that, for all $i \in [1;h], (f_i, g_i)$ is free for $X \backslash \{f_1, g_1, ..., f_{i-1}, g_{i-1}\}$ and $(X \backslash \{f_1, g_1, ..., f_k, g_k\})$ is a collapse of $X$.

Any $n$-face $c \in Y$ is such that $c \in X$ and $c \notin S$, and therefore there exists $j \in [1;h]$ such that $c = f_j$. Without loss of generality, let $j$ be the smallest integer such that $f_j \in Y$: for all $k \in [1;j-1]$, $f_k \notin Y$ and $g_k \notin Y$. As previously said, $(f_j, g_j)$ is free for $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$: $g_j$ is included in only one face of $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$, and this face is $f_j$.

As $\{f_1, g_1, ..., f_{j-1}, g_{j-1}\} \cap Y = \emptyset$ and that $f_j \in Y$, then $g_j$ is included in only one face of $Y$, and that face is $f_j$. Consequently, the pair $(f_j, g_j)$ is free for $Y$. Moreover, the pair $(f_j, g_j)$ belongs to the sequence $C$, therefore $f_j \notin S$ and $g_j \notin S$. As $W \subseteq S$, the pair $(f_j, g_j)$ is included in $Y \backslash W$.$\finproof$
\end{proof}

>>>>

In the 3d voxel framework, objects are mae of voxels. In the 3d cubical complex framework, objects made of voxel, but of cubes, squares, lines and vertices. 
Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$. Any subset $f$ of $\Z^n$ such that $f$ is the cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$. A $0$-face is called a \myem{vertex}, a $1$-face is an \myem{edge}, a $2$-face is a \myem{square}, and a $3$-face is a \myem{cube}.

@@

Informally, in 3d, a set of faces $X$ is each side (square) of a cube belonging to $X$ also belong to $X$, each edge of a square belonging to $X$ also belong to $X$, and each vertex of an edge belonging to $X$ also belong to $X$.

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. We denote by $\princ{X}$ the set composed of all facets of $X$ . A complex $X$ is \myem{pure} if all its facets have the same dimension. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.
In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$; in $\allfaces{3}$, a complex is thin if it contains no cube.

@@

Traditionally, a voxel image is defined as a finite subset of $\Zset^n$. This kind of image is the most common one in the field of image processing and we give now a simple way to transpose a binary image to the cubical complex framework.

Informally, to do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$ (to a pixel we associate a square, to a voxel we associate a cube). More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$. Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$.

@@

In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$. It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.

@@

If $Y$ is a collapse of a complex $X$, then $Y$ is a complex. 
We now introduce some elements that will serve for proving the thinness of our skeletons.
Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
(f_0,...,f_\ell)$ of faces of $\allfaces{n}$ such that for all $i \in \dinterval{0}{\ell}$, either $i$ is even and $f_i$ is an $n$-face, or $i$ is odd and $f_i$ is an $(n-1)$-face with $\contstrict{f_i}=\{f_{i-1}, f_{i+1}\}$ (such path always exists).

@@
